### 2023-11-26 documents/sections/extensions.tex
yukpa 1-say

\pex<yukpatr>\yukpa \parencite[][139]{meira2006syntactic}
\a<yuk-5>
\begingl
\gla aw j-esare//
\glb \gl{1}\gl{pro} \gl{3}>\gl{1}-see//
\glft \qu{S/he saw me.}//
\endgl
\a<yuk-1>
\begingl
\gla aw {\normalfont ∅}-esare//
\glb \gl{1}\gl{pro} \gl{1}>\gl{3}-see//
\glft \qu{I saw it.}//
\endgl
\xe
%
Since \PC \gl{1}\gl{sa} \rc{w(ɨ)-} was extended to \gl{1}>\gl{3} scenarios in most languages \parencite[81--82]{gildea1998}, and given its inclination for phonological erosion \pcref{sec:pekodian,sec:waiwaian}, the zero marking in \gl{1}>\gl{3} scenarios can be identified as the \yukpa reflex of \gl{1}\gl{sa} \rc{w-}.

Zero-marked first person is attested for a single intransitive verb, \obj{to} \qu{to go} \exref{yuk-7}.
It diverges from regular C-initial verbs with \obj{jɨ-}, like \qu{to sleep} \pcref{tab:yukreg}.
It can thus be identified as having resisted the extension of \obj{j-} in \yukpa.

\ex<yuk-7>\yukpa \parencite[][139]{meira2006syntactic}\\
\begingl
\gla aw {\normalfont ∅}-to//
\glb \gl{1}\gl{pro} \gl{1}\gl{sa}-go//
\glft \qu{I went.}//
\endgl
\xe

>>>>

\pex<yukpatr>\yukpa \parencite[][139]{meira2006syntactic}
\a<yuk-5>
\begingl
\gla aw j-esare//
\glb \gl{1}\gl{pro} \gl{3}>\gl{1}-see//
\glft \qu{S/he saw me.}//
\endgl
\a<yuk-1>
\begingl
\gla aw {\normalfont ∅}-esare//
\glb \gl{1}\gl{pro} \gl{1}>\gl{3}-see//
\glft \qu{I saw it.}//
\endgl
\xe
%
Since \PC \gl{1}\gl{sa} \rc{w(ɨ)-} was extended to \gl{1}>\gl{3} scenarios in most languages \parencite[81--82]{gildea1998}, and given its inclination for phonological erosion \pcref{sec:pekodian,sec:waiwaian}, the zero marking in \gl{1}>\gl{3} scenarios can be identified as the \yukpa reflex of \gl{1}\gl{sa} \rc{w-}.
Zero-marked first person is attested for a two intransitive verbs, \obj{to} \qu{to go} and \obj{ka} \qu{to say} \exref{yuk-zero}.
They diverge from regular C-initial verbs with \obj{jɨ-}, like \qu{to sleep} \pcref{tab:yuk-reg} and can thus be identified as having resisted the extension of \obj{j-} in \yukpa.

\pex<yuk-zero> \yukpa \parencites[139]{meira2006syntactic}{meiraDByukpa}
\a<yuk-7>
\begingl
\gla aw {\normalfont ∅}-to//
\glb \gl{1}\gl{pro} \gl{1}\gl{sa}-go//
\glft \qu{I went.}//
\endgl
\a<yuk-say>
\begingl
\gla awɨ sa {\normalfont ∅}-ka jɨ-na//
\glb \gl{1}\gl{pro} thus \gl{1}\gl{sa}-say \gl{3}-\gl{obl}//
\glft \qu{I said to him,}//
\endgl
\xe
